---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chinautils

<!-- badges: start -->
<!-- badges: end -->

This package provides a collection of helper functions to make working with Chinese (administrative) data easier.

## Installation

You can install the development version of chinautils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("turbanisch/chinautils")
```

## Harmonize province codes 

The names of Chinese provinces may vary between institutions and data sources: Inner Mongolia might come as "Inner Mongol", "Inner Mongolia Autonomous Region", "内蒙古", "内蒙古自治区" or "Nei Menggu Zizhiqu", making it hard to merge province-level data. `provincename` uses regular expressions to convert province names in English, German or Chinese into ISO codes and other output formats, such as those used by China Customs. 

```{r example}
library(chinautils)

provincename("Innere Mongolei")
provincename("Hong Kong", destination = "full_name_zh")
provincename("Formosa", destination = "type")
```

