---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  suppressPackageStartupMessages(library(tidyverse))
)
```

```{r, include=FALSE}
library(chinautils)
```

# chinautils

<!-- badges: start -->
<!-- badges: end -->

This package contains helper functions to make working with Chinese (administrative) data easier.

## Installation

You can install the development version of chinautils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("turbanisch/chinautils")
```

## Harmonize province codes 

The names of Chinese provinces may vary between data sources: Inner Mongolia might come as "Inner Mongol", "Inner Mongolia Autonomous Region", "内蒙古", "内蒙古自治区", "Nei Menggu" or "Nei Menggu Zizhiqu", making it hard to merge province-level data. 

`provincename` uses regular expressions to convert province names in English, German or Chinese to ISO codes and other output formats, such as those used by China Customs. 

```{r example}
provincename("Innere Mongolei")
provincename("Hong Kong", destination = "full_name_zh")
provincename("黑龙江", destination = "china_customs")
```

The function does not, however, take care of spelling errors that are pervasive in Chinese administrative data. Usually, the name in Chinese characters is the least error-prone. When regular expressions fail to capture province names, fuzzy matching using string distance based on the custom dictionary `chinautils::province_dict` will likely get the job done:

```{r}
library(tidyverse)

# create (misspelled) province names to be matched
df <- tibble(
  misspelled = c("Schanghai",
                 "Peking",
                 "Innere Mongolei", 
                 "Hong Kong", 
                 "Hong Kong Special Administrative Zone",
                 "Xingjiang")
)

# filter cartesian join by shortest distance
df %>% 
  left_join(select(province_dict, short_name_en),
            by = character(0)) %>% 
  mutate(dist = stringdist::stringdist(misspelled, short_name_en)) %>% 
  group_by(misspelled) %>% 
  filter(min_rank(dist) == 1L) %>% 
  ungroup()
```

