---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  suppressPackageStartupMessages(library(tidyverse))
)
```

```{r, include=FALSE}
library(chinautils)
```

# chinautils

<!-- badges: start -->
<!-- badges: end -->

This package contains helper functions to make working with Chinese (administrative) data easier.

## Installation

You can install the development version of chinautils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("turbanisch/chinautils")
```

## Harmonize country names (in Chinese)

Coming up with regular expressions to match country names in Chinese is slightly more involved than for other languages. The reason is that different parts of the world use different variants of Chinese, beyond the basic distinction between simplified and traditional characters. Mainland China, Malaysia and Singapore all use simplified characters whereas Hong Kong, Macau and Taiwan continue to use traditional characters -- but local usage may vary within each group. 

This is especially true for proper names like country names. Not only can they vary character by character depending on the script that is used, but they might also reflect different (phonetic) transliterations or refer to another name altogether. Here are some examples for each case:

1. **Different scripts.** Germany is referred to as *Deguo* 德国 in Mainland China and 德國 in Taiwan -- where 国 is the simplified character corresponding to 國, easy. However, there are some less obvious cases, as we will see below.
2. **Different transliterations**. Many country names have been phonetically adapted from other languages and translators in every Chinese-speaking region have taken their artistic liberties when doing so. For example, Hong Kongers refer to Barbados as *Babaduosi* 巴巴多斯 whereas people from Taiwan call it *Babeiduo* 巴貝多.
3. **Alternative names**. Instead of adapting a country name according to its sound, Chinese-speaking people in some regions have also opted to convert the original meaning of the country name into Chinese. For example, Montenegro is *Heishan* 黑山 ("black mountain") in Mainland China. People on Taiwan, on the other hand, kept their phonetic transliteration *Mengteneigeluo* 蒙特內哥羅.

The function `countryname` identifies country names in Chinese and converts them to various standardized output formats, such as ISO3 codes. 
It uses regular expressions to match country name variants in both simplified and traditional Chinese:
    
```{r}
# match variants in both simplified and traditional Chinese
countryname(c("中国", "中华人民共和国", "亞東開化中國早"))

# regex ignore languages other than Chinese and ambiguous cases
countryname(c("ドイツ国", "刚果"))

# get warned about potential pitfalls, such as multiple matches
countryname(c("塞尔维亚和黑山", "捷克斯洛伐克", "德国德国"))

# non-regex matching requires an exact match
countryname(c("德国", "德国人"), origin = "short_name_zh_cn", destination = "short_name_en")
```

Methodology and code for the conversion table can be found in [this repo](https://github.com/turbanisch/chinese-countryname-regex).

## Harmonize province codes 

The names of Chinese provinces may vary between data sources: Inner Mongolia might come as "Inner Mongol", "Inner Mongolia Autonomous Region", "内蒙古", "内蒙古自治区", "Nei Menggu" or "Nei Menggu Zizhiqu", making it hard to merge province-level data. 

`provincename` uses regular expressions to convert province names in English, German or Chinese to ISO codes and other output formats, such as those used by China Customs. 

```{r example}
provincename("Innere Mongolei")
provincename("Hong Kong", destination = "full_name_zh")
provincename("黑龙江", destination = "china_customs")
```

The function does not, however, take care of spelling errors that are pervasive in Chinese administrative data. Usually, the name in Chinese characters is the least error-prone. When regular expressions fail to capture province names, fuzzy matching using string distance based on the custom dictionary `chinautils::province_dict` will likely get the job done:

```{r}
library(tidyverse)

# create (misspelled) province names to be matched
df <- tibble(
  misspelled = c("Schanghai",
                 "Peking",
                 "Innere Mongolei", 
                 "Hong Kong", 
                 "Hong Kong Special Administrative Zone",
                 "Xingjiang")
)

# filter cartesian join by shortest distance
df %>% 
  left_join(select(province_dict, short_name_en),
            by = character(0)) %>% 
  mutate(dist = stringdist::stringdist(misspelled, short_name_en)) %>% 
  group_by(misspelled) %>% 
  filter(min_rank(dist) == 1L) %>% 
  ungroup()
```

