---
title: "Reading China Customs trade data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading China Customs trade data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(chinautils)
library(dplyr)
```

The [China Customs statistics portal](http://stats.customs.gov.cn/indexEn)
lets you download trade data as CSV files. These files are awkward to read
programmatically:

- they are encoded as **GB18030**, not UTF-8;
- column names come in **English or Chinese** depending on the interface
  language, and you may end up with a mix if you download several queries;
- the **set and order of columns depends on the query** -- a single-month
  download has no date column, and the sort order changes the column order.

`cc_read_csv()` hides all of this. It reads one or more files into a single
tidy tibble with stable English column names.

## A single file

This package ships a few small example downloads. A "full" English query
contains the date, the trade dimensions (commodity, partner, customs regime,
province) and the value and quantity columns:

```{r}
path <- system.file("extdata", "english-full.csv", package = "chinautils")
cc_read_csv(path)
```

Codes are kept as character vectors (so leading zeros survive), the date is
parsed, and the value column is numeric.

## Many files at once, in any language

Pass a vector of paths to read and row-bind several files -- even if they are
in different languages and have different columns. Here we combine an English
file (with a month column) and a Chinese file (a multi-month aggregate, hence
no month column):

```{r}
files <- system.file(
  "extdata",
  c("english-full.csv", "chinese-full.csv"),
  package = "chinautils"
)
cc_read_csv(files)
```

The Chinese rows have a missing `yearmonth` and populate `value_cny` instead of
`value_usd`: values reported in different currencies are deliberately kept in
separate columns so they are never accidentally summed.

## Single-month downloads

If you download a single month, China Customs omits the date column entirely.
`cc_read_csv()` handles this without complaint -- it simply returns a table
without `yearmonth`:

```{r}
degen <- system.file("extdata", "english-degenerate.csv", package = "chinautils")
cc_read_csv(degen)
```

## Re-attaching clean descriptions

By default `cc_read_csv()` drops the descriptions that ship alongside each code,
because they are redundant and frequently misspelled. Re-attach clean
descriptions from the bundled lookup tables when you need them:

```{r}
trade <- cc_read_csv(path)

trade |>
  left_join(cc_partners, by = c("partner" = "code")) |>
  select(partner, partner_en = en, value_usd)
```

Commodity descriptions are versioned by year, so join `cc_commodities` on both
the commodity code and the year:

```{r}
trade |>
  mutate(year = lubridate::year(yearmonth)) |>
  left_join(cc_commodities, by = c("commodity" = "code", "year")) |>
  distinct(commodity, year, en)
```

## Things to watch out for

`cc_read_csv()` warns about several silent pitfalls of the download page:

- **Truncation.** The portal caps a download at 10,000 rows. If a file has
  exactly 10,000 data rows it may be incomplete, and you get a warning.
- **Encoding.** If the column names cannot be decoded as GB18030 (for example
  because a file was re-saved as UTF-8), `cc_read_csv()` stops with a clear
  message; pass `encoding = "UTF-8"` to override the default.
- **Mismatched columns.** When you read several files at once and they do not
  all share the same columns, you get a warning, because the affected rows are
  filled with `NA`. This is the case described above when you append a
  single-month update to a dated series.

One pitfall cannot be detected automatically: the **trade flow**. Imports,
exports and combined "import and export" downloads all have exactly the same
columns -- the direction is never stored in the file. If you read an
imports-only file together with an exports-only file, the two are merged without
warning. Tag each file yourself before binding if you need to keep them apart:

```{r eval = FALSE}
imports <- cc_read_csv("imports.csv") |> mutate(flow = "import")
exports <- cc_read_csv("exports.csv") |> mutate(flow = "export")
bind_rows(imports, exports)
```
